Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention

### 2024-06-25 manuscripts/manuscript1.qmd
prior update

---
title: Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention
author: "Nikhil Patel, MD"
subtitle: A Bayesian Re-Analysis of ATACH-2 and ERICH data
authors:
  - name: Nikhil Patel
    orcid: 0000-0002-3193-8590
    affiliation: Department of Internal Medicine, Wake Forest University School of Medicine
    roles: data analysis, writing
    corresponding: true
  - name: Matt Jaffa
    affiliation: Department of Neurology, University of Connecticut School of Medicine
  - name: Rahul Karamchandani
    affiliation: Department of Neurology, Wake Forest University School of Medicine
  - name: Andrew Asimos
    affiliation: Department of Emergency Medicine, Wake Forest University School of Medicine
  - name: Jonathan Parish
    affiliation: Carolina Neurosurgery & Spine Associates
  - name: Nicholas Allen Morris
    affiliation: Department of Neurology, University of Maryland School of Medicine
date: last-modified
abstract: |

@@

#### Covariates
We constructed directed acyclic graphs (DAGs) to model our causal assumptions and to appropriately select variables to minimize confounding and selection bias. Based on our specifications, we included the following covariates in our models: age, admission Glasgow coma score (GCS), ICH location (categorized as lobar, thalamic, or basal ganglia), baseline ICH volume (measured in millilters), and presence of intraventricular hemorrhage (IVH). The DAGs used in the analyses are available in the supplementary material.

#### Priors
Two sets of priors were applied for all models, both assuming no association of hemispheric laterality on the outcomes of interest: 1) default vague priors as defined by the 'brms' software package; 2) flat priors; 3) weakly informative priors based on previous literature. A description of the reasons for these priors is available in the supplementary material.

#### Posterior distribution
We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median odds ratio (OR) of the posterior distribution with 95% credible interval; 2) probability of any difference (P[OR] > 1); 3) probability of a large difference (P[OR] > 1.2) as defined by the study authors; 4) percent of the posterior distribution outside of the region of practical equivalence as defined by the study authors (ROPE; 0.95 > P[OR] > 1.05). A description of the reasons for these thresholds is available in the supplementary material.

>>>>

---
title: Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention
author: "Nikhil Patel, MD"
subtitle: A Post Hoc Bayesian of ATACH-2 and ERICH data
authors:
  - name: Nikhil Patel
    orcid: 0000-0002-3193-8590
    affiliation: Department of Internal Medicine, Wake Forest University School of Medicine
    roles: data analysis, writing
    corresponding: true
  - name: Matt Jaffa
    affiliation: Department of Neurology, University of Connecticut School of Medicine
  - name: Rahul Karamchandani
    affiliation: Department of Neurology, Wake Forest University School of Medicine
  - name: Andrew Asimos
    affiliation: Department of Emergency Medicine, Wake Forest University School of Medicine
  - name: Jonathan Parish
    affiliation: Carolina Neurosurgery & Spine Associates
  - name: Nicholas Allen Morris
    affiliation: Department of Neurology, University of Maryland School of Medicine
date: last-modified
abstract: |

@@

#### Covariate Selection
We constructed directed acyclic graphs (DAGs) to model our causal assumptions and to appropriately select variables to minimize confounding and selection bias. Based on our specifications, we included the following covariates in our models: age, admission Glasgow coma score (GCS), ICH location (categorized as lobar, thalamic, or basal ganglia), baseline ICH volume (measured in millilters), and presence of intraventricular hemorrhage (IVH). The DAGs used in the analyses are available in the supplementary material.

#### Prior Distribution Specification
Two sets of priors were applied for all models, both assuming no association of hemispheric laterality on the outcomes of interest: 1) default, weakly informative priors as defined by the 'brms' software package; 2) priors informed from review of previous literature. A description of the reasons for these priors is available in the supplementary material.

#### Computation and Description of the Posterior Distribution
We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median odds ratio (OR) of the posterior distribution with 95% credible interval; 2) probability of any difference (P[OR] > 1); 3) probability of a large difference (P[OR] > 1.2) as defined by the study authors; 4) percent of the posterior distribution outside of the region of practical equivalence as defined by the study authors (ROPE; 0.95 > P[OR] > 1.05). A description of the reasons for these thresholds is available in the supplementary material.
